dispRity palaeo demo

### 2017-06-19 doc/dispRity-palaeo-demo.Rmd
Revert "Merge pull request #33 from TGuillerme/master"
This reverts commit 53d9a8f98c5ad8d65cc1f4466045e3da67ef2d8d, reversing
changes made to dac6e1109659bb50d7f261c03a9e0969fe791807.

One of the crucial steps in disparity-through-time analysis is to split the full morphospace into smaller time subsamples that contain the total amount of morphologies at certain points in time (time-slicing) or during certain periods in time (time-binning).
Basically, the full morphospace in the previous step represents the total amount of morphologies in all time and is likely to be greater than any time subsamples of the morphospace.

The `dispRity` package provides a `time.subsamples` function that allows to split the morphospace time slices (using `method = continuous`) or into time bins (using `method = discrete`).
In this example, we are going to split the morphospace into 5 equal time bins of 20 million years long from 100 million years ago (Mya) to the present.
We will also provide to the function the table containing the first and last occurrences data for some fossils to take into account the fact that some fossils might span between our different time bins.

@@

## Splitting the morphospace using the time.subsamples function
(binned_morphospace <- time.subsamples(data = BeckLee_mat50, tree = BeckLee_tree,
    method = "discrete", time = time_bins, inc.nodes = FALSE,
    FADLAD = BeckLee_ages))
```

The output object is a `dispRity` object.
For details about this object class, please refer to the [`dispRity` manual](http://htmlpreview.github.com/?https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-manual.html).
In brief, however, `dispRity` objects are lists of different elements (i.e. disparity results, morphospace time subsamples, morphospace attributes, etc.) that display only a summary of the object when calling the object for avoiding jamming the `R` console.

@@

Once we obtain our different time subsamples, we can bootstrap and rarefy them (i.e. pseudo-replicating the data).
The bootstrapping will allow to make each subsamples more robust to outliers and the rarefaction will allow us to compare subsamples with the same number of taxa to get rid of eventual sampling problems (i.e. more taxa in one subsamples than the other).
The `boot.matrix` function that allows to bootstrap the `dispRity` object and the `rarefaction` option within allows to rarefy it.

```{r}
## Bootstrapping each time subsamples 100 times
(boot_bin_morphospace <- boot.matrix(binned_morphospace, bootstraps = 100))

## Getting the minimum number of rows (=taxa) in the time subsamples
min(unlist(lapply(boot_bin_morphospace$subsamples, lapply, nrow)))

## Bootstrapping each time subsamples 100 times and rarefying them 
(rare_bin_morphospace <- boot.matrix(binned_morphospace, bootstraps = 100,
    rarefaction = 6))
```
Calculating disparity
---------------------

We can now calculate the disparity within each time subsamples along side with some confidence intervals generated by the pseudo-replication step above (bootstraps/rarefaction).
Disparity can be calculated in many ways and this package allows user to come up with their own disparity metrics.
For more info, please refer to the [`dispRity` metric vignette](http://htmlpreview.github.com/?https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-metrics.html).

In this example, we are going to calculate the spread of the data in each time subsamples by calculating the disparity as the sum of the variance of each dimension of the morphospace in each time subsamples using the `dispRity` function.
Thus, in this example, the disparity is defined by the multi-dimensional variance of each time subsamples (i.e. the spread of the taxa within the morphospace).
Note that this metric comes with a caveat not solved here since it ignores the covariance among the dimensions of the morphospace.

@@

The `dispRity` function does not actually display the calculated disparity values but rather only the properties of the disparity object (size, subsamples, metric, etc.).
To display the actual calculated scores, we need to summarise the disparity object using the S3 method `summary` that is applied on a `dispRity` object (see `?summary.dispRity` for more details).

@@

Finally, to draw some valid conclusions from these results, we can apply some statistical testing to the data.
We can test, for example, if mammalian disparity changed through time since the last 100 million years (Mya).
To do so, we can apply some comparisons of the means between each time-bins in a sequential manner to see whether the disparity in bin *n* is equal to the disparity in bin *n+1* which is in turn equal to the disparity in bin *n+2*, etc.
Because our data is temporally auto-correlated (i.e. what happens in bin *n+1* depends on what happened in bin *n*) and was pseudo-replicated (i.e. each bootstrap draw creates non-independent time subsamples because they are all based on the same time subsamples), we are going to apply a non-parametric mean comparison: the `wilcox.test`.
Also, we are going to apply a p-value correction to avoid type I error inflation due to performing multiple test for the same hypothesis (see `?p.adjust` for more details).

>>>>

One of the crucial steps in disparity-through-time analysis is to split the full morphospace into smaller time series that contain the total amount of morphologies at certain points in time (time-slicing) or during certain periods in time (time-binning).
Basically, the full morphospace in the previous step represents the total amount of morphologies in all time and is likely to be greater than any time series of the morphospace.

The `dispRity` package provides a `time.series` function that allows to split the morphospace time slices (using `method = continuous`) or into time bins (using `method = discrete`).
In this example, we are going to split the morphospace into 5 equal time bins of 20 million years long from 100 million years ago (Mya) to the present.
We will also provide to the function the table containing the first and last occurrences data for some fossils to take into account the fact that some fossils might span between our different time bins.

@@

## Splitting the morphospace using the time.series function
(binned_morphospace <- time.series(data = BeckLee_mat50, tree = BeckLee_tree,
    method = "discrete", time = time_bins, inc.nodes = FALSE,
    FADLAD = BeckLee_ages))
```

The output object is a `dispRity` object.
For details about this object class, please refer to the [`dispRity` manual](http://htmlpreview.github.com/?https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-manual.html).
In brief, however, `dispRity` objects are lists of different elements (i.e. disparity results, morphospace time series, morphospace attributes, etc.) that display only a summary of the object when calling the object for avoiding jamming the `R` console.

@@

Once we obtain our different time series, we can bootstrap and rarefy them (i.e. pseudo-replicating the data).
The bootstrapping will allow to make each series more robust to outliers and the rarefaction will allow us to compare series with the same number of taxa to get rid of eventual sampling problems (i.e. more taxa in one series than the other).
The `boot.matrix` function that allows to bootstrap the `dispRity` object and the `rarefaction` option within allows to rarefy it.

```{r}
## Bootstrapping each time series 100 times
(boot_bin_morphospace <- boot.matrix(binned_morphospace, bootstraps = 100))

## Getting the minimum number of rows (=taxa) in the time series
min(unlist(lapply(boot_bin_morphospace$series, lapply, nrow)))

## Bootstrapping each time series 100 times and rarefying them 
(rare_bin_morphospace <- boot.matrix(binned_morphospace, bootstraps = 100,
    rarefaction = 6))
```
Calculating disparity
---------------------

We can now calculate the disparity within each time series along side with some confidence intervals generated by the pseudo-replication step above (bootstraps/rarefaction).
Disparity can be calculated in many ways and this package allows user to come up with their own disparity metrics.
For more info, please refer to the [`dispRity` metric vignette](http://htmlpreview.github.com/?https://github.com/TGuillerme/dispRity/blob/master/doc/dispRity-metrics.html).

In this example, we are going to calculate the spread of the data in each time series by calculating the disparity as the sum of the variance of each dimension of the morphospace in each time series using the `dispRity` function.
Thus, in this example, the disparity is defined by the multi-dimensional variance of each time series (i.e. the spread of the taxa within the morphospace).
Note that this metric comes with a caveat not solved here since it ignores the covariance among the dimensions of the morphospace.

@@

The `dispRity` function does not actually display the calculated disparity values but rather only the properties of the disparity object (size, series, metric, etc.).
To display the actual calculated scores, we need to summarise the disparity object using the S3 method `summary` that is applied on a `dispRity` object (see `?summary.dispRity` for more details).

@@

Finally, to draw some valid conclusions from these results, we can apply some statistical testing to the data.
We can test, for example, if mammalian disparity changed through time since the last 100 million years (Mya).
To do so, we can apply some comparisons of the means between each time-bins in a sequential manner to see whether the disparity in bin *n* is equal to the disparity in bin *n+1* which is in turn equal to the disparity in bin *n+2*, etc.
Because our data is temporally auto-correlated (i.e. what happens in bin *n+1* depends on what happened in bin *n*) and was pseudo-replicated (i.e. each bootstrap draw creates non-independent time series because they are all based on the same time series), we are going to apply a non-parametric mean comparison: the `wilcox.test`.
Also, we are going to apply a p-value correction to avoid type I error inflation due to performing multiple test for the same hypothesis (see `?p.adjust` for more details).
